---
title: "Initial Korean covid 19"
author: "Sophia Castellon"
date: "October 16 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Methods

$H_0$ There is no difference in dying depending on sex
$H_A$ There is a difference in dying depending on sex  

This data is from Korea in 2022

Result~sex
Cat~Cat helper file
  
# Your Question - Is dying independent of sex?

## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

This bar chart shows that more males died the women. 

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Among those who died there was 64% of men who died and 36% of women who died.  

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi Squared test shows the p value is 0.001 and the Fisher Exact test  has a p value of 0.00009. The odds of the Fisher Exact test  states that men are 2.3 more acceptable of dying from covid 19. Since both p values are lower then 0.05 we reject the null hypothesis.  

## Conclusions

Is dying independent of sex?

$H_0$ There is no difference in dying depending on sex
$H_A$ There is a difference in dying depending on sex  

This data is from Korea in 2022

We reject the null hypothesis for the following,  
The bar chart shows that more males died then women.  
The numerical data shows the chi squared p value is 0.001 and the Fisher Exact data shows the p value was 0.00009. The odd of the Fisher Exact test data shows that men are 2.3 more acceptable of dying from covid 19. Therefore since both p vales are lower then 0.05 we reject the null hypothesis.  


